### 2015-06-02 mdp.tex
Cosmetics and some comments, especially on the proof of Lemma 8.4

Given an SDP $u$, the infinite horizon discounted reward corresponding to state $s$ under $u$ is denoted by $J_u(s)$ and is defined by
\begin{align}
J_u(s)\stackrel{\Delta}{=}\E[\sum_{n=0}^\infty \alpha^n g_{a_n}(s_n)|s_0=s,a_n=u(s_n), n= 0,1,\dots],\nn
\end{align}
where $\alpha \in (0,1)$ is a given discount factor. Here $J_u(s)$ is known as the value of the state $s$ under the SDP $u$, and the vector quantity $J_u\stackrel{\Delta}{=}(J_u(s))_{s\in S}\in R^n$ is called the value-function corresponding to the SDP $u$.\\
The optimal policy $u^*$ is any policy whose value is the largest possible under any policy: 
$J_{u^*}(s) = \max_{u\in U} J_u(s)$ for all $s\in S$. Such $u^*$ exists and is well defined in the case of infinite horizon discounted reward MDP \cite{Puter}.
\todoc{A stronger definition allows any policy, including history dependent and yet the statement will hold true.}
 The optimal value-function $J^*$ is the one obtained under the optimal policy, i.e., $J^*=J_{u^*}$. \todoc{Probably better to define first $J^*$ as the pointwise max and then the optimal policy's definition could use $J^*$.}
Given an MDP, our aim is to find the optimal value function $J^*$ and the optimal policy $u^*$. The optimal policy and value function obey the Bellman equation (BE): for all $ s \in S$, \todoc{Non-unique maximizers in the policy def?}
\begin{subequations}\label{bell}
\begin{align}
\label{bellval}J^*(s)&=\max_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big),\\
\label{bellpol}u^*(s)&=\arg\max_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big).
\end{align}
\end{subequations}
If $J^*$ is computed first, then $u^*$ can be obtained by substituting $J^*$ in \eqref{bellpol}. The Bellman operator $T$ is defined using the model parameters of the MDP as follows:
\begin{definition}
The Bellman operator $T: \R^n \to \R^n$ is defined by 
\begin{align}
(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big)\,, 
\end{align}
where $J\in \R^n$ and $J_s = J(s)$, $s\in S = \{1,\dots,n\}$.  
Similarly one can define the Bellman operator $T_u:\R^n \to \R^n$ restricted to an SDP $u$ as follows:
\begin{align}
(T_uJ)(s)=g_{u(s)}(s)+\alpha \sum_{s'} p_{u(s)}(s,s')J(s')\,.
\end{align}
\end{definition}
Given $J \in \R^n$, $TJ$ is the `one-step' greedy value function. It is also useful to define the notion of a one-step greedy policy as below:
\begin{definition}
A policy $\tilde{u}$ is said to be greedy with respect to $\tilde{J}$ if
\begin{align}\label{subpol}
\tilde{u}(s)=\underset{a \in A}{\arg\max}(g_a(s)+\alpha\sum_{s'} p_a(s,s')\tilde{J}(s')).
\end{align}
\end{definition}
\begin{definition}
We also define the $Q$-Bellman operator \cite{BertB} $H$. Given any $J\in \R^n$ we have,

\begin{align}
HJ&=\left[ {\begin{array}{c} H_1 J  \\ \vdots \\ H_d J\end{array}} \right]\in \R^{nd},\mbox{where}\nn\\
(H_a J)(s)&= g_a(s)+\alpha \sum_{s'}p_a(s,s') J(s'), \qquad s\in S, a\in A,
\end{align}
\end{definition}
 We now state without proof the important properties related to the Bellman operator. Though in these results, we make use of the Bellman operator $T$, the results also trivially hold for $T_u$ as well.
\subsection{Properties of $T$}
\begin{lemma}\label{maxnorm}
$T$ is a $\max$-norm contraction operator, i.e., given $J_1, J_2 \in \mathbf{R}^n$,
\begin{align}
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
\end{align}
\end{lemma}
\begin{corollary}\label{uniquesol}
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{corollary}
The Bellman operator $T$ exhibits two more important properties presented in the following lemmas (see \cite{BertB} for proofs):
\begin{lemma}\label{monotone}
$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_2\geq J_1$, we have $T J_2\geq T J_1$. Further, if $J\in \R^n$ is such that $J\geq TJ$, it follows that $J\geq J^*$.
\end{lemma}
\begin{lemma}\label{shift}
Given $J\in \R^n$, $t \in \R$ and $\mathbf{1} \in \R^n$, a vector with all entries $1$, we have
\begin{align}
T(J+t\mathbf{1})=TJ+\alpha t\mathbf{1}.
\end{align}
\end{lemma}
Note that Lemmas~\ref{maxnorm}-\ref{shift} also hold for the $Q$-Bellman operator $H$.
Solving an MDP involves handling two sub-problems namely the problem of \emph{control} and the problem of \emph{prediction}. The problem of \emph{control} deals with coming up with a good (and if possible the optimal) policy. Often, in order to solve the problem of \emph{control}, one needs to solve the problem of \emph{prediction}, which deals with computing the value function $J_u$ of the policy $u$. The fact that the two problems are related is reflected in the Bellman equation in \eqref{bell}, where $J^*$ from \eqref{bellval} is used in \eqref{bellpol} to obtain the optimal policy $u^*$. Thus, the Bellman equation is at the heart of the solution methods to MDPs. Any solution method to MDP is said to be complete only if it satisfactorily (with provable performance guarantees) addresses both the prediction and the control problems. \\
The basic solution methods namely value iteration, policy iteration and linear programming (LP) formulation \cite{BertB} solve both the control and prediction problems. Of the three basic methods, in this paper, we are interested in the LP formulation given by
\begin{align}\label{mdplp}
\min_{J\in \R^n} &c^\top J\nn\\
\text{s.t}\mb &J(s)\geq g_a(s)+\alpha\sum_{s'}p_a(s,s')J(s'), \qquad s\in S, a \in A,
\end{align}
where $c\in \R^n_+$ is any vector whose components are all non-negative. One can show that $J^*$ is the solution to the LP formulation \eqref{mdplp} \cite{BertB}. 
Also, of the three methods, value iteration and LP formulation are value function based methods, i.e., they compute $J^*$ directly and then $u^*$ is obtained by plugging $J^*$ in \eqref{bellpol}.\\
While the basic methods (i.e., VI, PI and LP) can be used to compute exact values $J^*$ and $u^*$ for MDPs with a small number of states, they are computationally expensive in the case of MDPs with a large number of states.\\

>>>>

Given an SDP $u$, the infinite horizon discounted reward corresponding to state $s$ under $u$ is denoted by $J_u(s)$ and is defined by
\begin{align}
J_u(s)\stackrel{\Delta}{=}\E[\sum_{n=0}^\infty \alpha^n g_{a_n}(s_n)|s_0=s,a_n=u(s_n), n= 0,1,\dots],\nn
\end{align}
where $\alpha \in (0,1)$ is a given discount factor. Here $J_u(s)$ is known as the value of the state $s$ under the SDP $u$, and the vector quantity $J_u\stackrel{\Delta}{=}(J_u(s))_{s\in S}\in R^n$ is called the value-function corresponding to the SDP $u$.\\
The optimal policy $u^*$ is any policy whose value is the largest possible under any policy: 
$J_{u^*}(s) = \max_{u\in U} J_u(s)$ for all $s\in S$. Such $u^*$ exists and is well defined in the case of infinite horizon discounted reward MDP \cite{Puter}.
\todoc{A stronger definition allows any policy, including history dependent and yet the statement will hold true.}
 The optimal value-function $J^*$ is the one obtained under the optimal policy, i.e., $J^*=J_{u^*}$. \todoc{Probably better to define first $J^*$ as the pointwise max and then the optimal policy's definition could use $J^*$.}
Given an MDP, our aim is to find the optimal value function $J^*$ and the optimal policy $u^*$. The optimal policy and value function obey the Bellman equation (BE): for all $ s \in S$, \todoc{Non-unique maximizers in the policy def?}
\begin{subequations}\label{bell}
\begin{align}
\label{bellval}J^*(s)&=\max_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big),\\
\label{bellpol}u^*(s)&=\arg\max_{ a\in A}\big(g_a(s)+\alpha \sum_{s'}p_a(s,s')J^*(s')\big).
\end{align}
\end{subequations}
If $J^*$ is computed first, then $u^*$ can be obtained by substituting $J^*$ in \eqref{bellpol}. The Bellman operator $T$ is defined using the model parameters of the MDP as follows:
\begin{definition}
The Bellman operator $T: \R^n \to \R^n$ is defined by 
\begin{align}
(TJ)(s)=\max_{a \in A}\big(g_a(s)+\alpha \sum_{s'} p_a(s,s')J(s')\big)\,, 
\end{align}
where $J\in \R^n$ and $J_s = J(s)$, $s\in S = \{1,\dots,n\}$.  
Similarly one can define the Bellman operator $T_u:\R^n \to \R^n$ restricted to an SDP $u$ as follows:
\begin{align}
(T_uJ)(s)=g_{u(s)}(s)+\alpha \sum_{s'} p_{u(s)}(s,s')J(s')\,.
\end{align}
\end{definition}
Given $J \in \R^n$, $TJ$ is the `one-step' greedy value function. It is also useful to define the notion of a one-step greedy policy as below:
\begin{definition}
A policy $\tilde{u}$ is said to be greedy with respect to $\tilde{J}$ if
\begin{align}\label{subpol}
\tilde{u}(s)=\underset{a \in A}{\argmax} \, \big(g_a(s)+\alpha\sum_{s'} p_a(s,s')\tilde{J}(s')\big)\,.
\end{align}
\end{definition}
We also define the Bellman operator $H$ for action values  \cite{BertB}: \todoc{$Q$-Bellman is a bad name.}
\begin{definition}
Let $H: \R^n \to \R^{nd}$ be defined as follows: For $J\in \R^n$,
\begin{align}
HJ&=\left[ {\begin{array}{c} H_1 J  \\ \vdots \\ H_d J\end{array}} \right]\in \R^{nd},\mbox{where}\nn\\
(H_a J)(s)&= g_a(s)+\alpha \sum_{s'}p_a(s,s') J(s'), \qquad s\in S, a\in A.
\end{align}
\end{definition}
 We now state without proof the important properties related to the Bellman operator. Though in these results, we make use of the Bellman operator $T$, the results also trivially hold for $T_u$ as well.
\subsection{Properties of $T$}
\begin{lemma}\label{maxnorm}
$T$ is a $\max$-norm contraction operator, i.e., given $J_1, J_2 \in \mathbf{R}^n$,
\begin{align}
||TJ_1-TJ_2||_\infty\leq \alpha ||J_1-J_2||_\infty.
\end{align}
\end{lemma}
\begin{corollary}\label{uniquesol}
$J^*$ is a unique fixed point of $T$, i.e., $J^*=TJ^*$.
\end{corollary}
The Bellman operator $T$ exhibits two more important properties presented in the following lemmas (see \cite{BertB} for proofs):
\begin{lemma}\label{monotone}
$T$ is a monotone map, i.e., given $J_1,J_2 \in \R^n$ such that $J_2\geq J_1$, we have $T J_2\geq T J_1$. Further, if $J\in \R^n$ is such that $J\geq TJ$, it follows that $J\geq J^*$.
\end{lemma}
\begin{lemma}\label{shift}
Given $J\in \R^n$, $t \in \R$ and $\mathbf{1} \in \R^n$, a vector with all entries $1$, we have
\begin{align}
T(J+t\mathbf{1})=TJ+\alpha t\mathbf{1}.
\end{align}
\end{lemma}
Note that Lemmas~\ref{maxnorm}-\ref{shift} also hold for the $Q$-Bellman operator $H$.
Solving an MDP involves handling two sub-problems namely the problem of \emph{control} and the problem of \emph{prediction}. The problem of \emph{control} deals with coming up with a good (and if possible the optimal) policy. Often, in order to solve the problem of \emph{control}, one needs to solve the problem of \emph{prediction}, which deals with computing the value function $J_u$ of the policy $u$. The fact that the two problems are related is reflected in the Bellman equation in \eqref{bell}, where $J^*$ from \eqref{bellval} is used in \eqref{bellpol} to obtain the optimal policy $u^*$. Thus, the Bellman equation is at the heart of the solution methods to MDPs. Any solution method to MDP is said to be complete only if it satisfactorily (with provable performance guarantees) addresses both the prediction and the control problems. \\
The basic solution methods namely value iteration, policy iteration and linear programming (LP) formulation \cite{BertB} solve both the control and prediction problems. Of the three basic methods, in this paper, we are interested in the LP formulation given by
\begin{align}\label{mdplp}
\begin{split}
\min_{J\in \R^n}\, &c^\top J\\
\text{s.t}\mb &J(s)\geq g_a(s)+\alpha\sum_{s'}p_a(s,s')J(s'), \qquad s\in S, a \in A,
\end{split}
\end{align}
where $c\in \R^n_+$ is any vector whose components are all non-negative. One can show that $J^*$ is the solution to the LP formulation \eqref{mdplp} \cite{BertB}. 
Also, of the three methods, value iteration and LP formulation are value function based methods, i.e., they compute $J^*$ directly and then $u^*$ is obtained by plugging $J^*$ in \eqref{bellpol}.\\
While the basic methods (i.e., VI, PI and LP) can be used to compute exact values $J^*$ and $u^*$ for MDPs with a small number of states, they are computationally expensive in the case of MDPs with a large number of states.\\
